Hierarchical Dispatching and Overriding

### 2017-10-10 NewVersion/sections/Formalization.tex
S3

\section{Formalization}~\label{sec:formalization}
In this section, we present the formal model \MIM{} (\emph{\textbf{F}eatherweight \textbf{H}ierarchical \textbf{J}ava}), based on
Featherweight Java~\cite{Igarashi01FJ}. The syntax, typing rules,
small-step semantics and theorems for type soundness are presented.
\bruno{In general, please try to use latex paragraphs to add more
  structure and emphasize contents.}

\vspace{-2ex}
\subsection{Syntax}
The abstract syntax of \MIM{} interface declarations, method declarations, and expressions is given at Figure~\ref{fig:syntax}. The multiple
inheritance feature of \MIM{} is inspired by Java 8 interfaces, which support
method implementations via default methods. This feature is 
closely related to \emph{traits}. To demonstrate how
unintentional method conflicts are untangled in \MIM{}, we only focus on a small subset of the interface model. For example, all the methods declared
in an interface are either default methods or abstract methods. Default methods provide default implementations for methods. Abstract methods do not
have method body, and in fact they can be considered as normal methods that return nothing. Abstract methods can be overridden with future implementations.

%From this point we can view that we
%are actually modelling a class model that supports multiple
%inheritance. 
Object creation is represented by $\new I$. If any unimplemented abstract methods are defined in interface $I$ (or inherited from parents), then object creation fails. Fields and primitive types are not modelled in \MIM{}.

The metavariables $I, J, K$ range over interface names; $x$ ranges over variables; $m$ ranges over method names; $e$ ranges over expressions; and $M$ ranges over method declarations. Following Featherweight Java, we assume that the set of variables includes the special variable \kwthis, which cannot be used as the name of an argument to a method. We use the same
conventions as FJ; we write $\overline{I}$ as shorthand for a possibly empty sequence $I_1, ..., I_n$, which may be indexed by $I_i$; and write $\overline{M}$ as shorthand for $M_1 .. M_n$ (with no commas). We also abbreviate operations on pairs of sequences in the obvious way, writing $\overline{I} \; \overline{x}$ for $I_1 \; x_1, ..., I_n \; x_n$, where $n$ is the length of $\overline{I}$ and $\overline{x}$.

In order to achieve multiple inheritance, an interface can have a set of 
parent interfaces, where such a set can be empty. The interface declaration $\interface{I}{I}{M}$ introduces an interface named $I$ with parent interfaces $\overline{I}$ and a suite of methods $\overline{M}$. The methods of $I$ may either override methods that are already defined in $\overline{I}$ or add new functionality special to $I$, we will illustrate this in more detail later.

The method declaration $\method{I}{m}{I_x}{x}{J}{e}$ introduces a
method named $m$ with result type $I$, parameters $\overline{x}$ of
type $\overline{I_x}$ and the overriding target $J$. The body of the
method is the single statement $ \kwreturn \; e;$. We also allow
defining abstract methods with $\absmethod{I}{m}{I_x}{x}{J}$, which is
similar to a normal method but without a method body. Notably, we have introduced the
\kwoverride{} keyword to override an old implementation of the
method. If the interface that it overrides is exactly the enclosing
interface, then such a method is seen as ``originally defined''; otherwise
it is hierarchical overriding.
For simplicity, overloading is not modelled for methods, which
implies that we can uniquely identify a method by its name.\yanlin{the terminology of original methods should have been proposed in Overview.}
\bruno{Then please do so.}

Expressions can be standard constructs such as variables, method
invocation, or object creation, together with casts.
Again, to be concise, the casts are merely safe up-casts. Those cast-expressions
are in fact annotated expressions, where the annotation indicates its static type.
The coexistence of static and dynamic types is the key to hierarchical dispatch.
A value
``$(I)\new{J}$''
is the final result of multiple reduction steps evaluating an
expression.

@@

There are two typing rules for method invocation: \textsc{(T-Invk)}
and \textsc{(T-PathInvk)}. \textsc{(T-Invk)} is the typing judgement
for a normal method invocation. First it ensures that the method
receiver $e_0$ type checks to a type $I_0$. Then it checks the type of
the specific method, by using an auxiliary function $\mbody(m, I_0,
I_0)$, which is the function for looking up methods that we will
illustrate in more detail in Section~\ref{subsec:auxdefs}. The form of
\mbody{} is different to FJ\bruno{Use xspace to fix the spacing problem
  of macros like mbody}, which has two parameters. The first $I_0$
represents the dynamic type and the second, the static type. If the
result of \mbody is not \Undefined, it means that the method exists and the corresponding method parameters and return types are given. After the method type is obtained, we check that all arguments $\overline{e}$ type check to $\overline{I}$ which are subtypes of $\overline{J}$. \textsc{(T-PathInvk)} is the typing judgement for a path invocation. Besides the conditions of \textsc{(T-Invk)}, \textsc{(T-PathInvk)} requires the type of receiver to be the subtype of the specified path type. 
% and \textsc{(T-SuperInvk)} checks if the enclosing type directly extends the specified super type.

@@

\bruno{Explanation for T-Anno is missing.}\yanlin{Added.}
Finally, rule \textsc{(T-Anno)} is the typing rule for an expression with static type annotation. The rule says that an annotated expression $\angl{J}e$ type checks to $J$ if the expression $e$ type checks and the static annotation $J$ is a subtype of $I$ which is the type of $e$.

>>>>

\section{Formalization}~\label{sec:formalization}
In this section, we present the formal model called \MIM{} (\emph{\textbf{F}eatherweight \textbf{H}ierarchical \textbf{J}ava}), following similar style of 
Featherweight Java~\cite{Igarashi01FJ}. The syntax, typing rules,
small-step semantics and theorems for type soundness are presented.

% \vspace{-2ex}
\subsection{Syntax}
The abstract syntax of \MIM{} interface declarations, method declarations, and expressions is given at Figure~\ref{fig:syntax}. The multiple
inheritance feature of \MIM{} is inspired by Java 8 interfaces, which support
method implementations via default methods. This feature is 
closely related to \emph{traits}. To demonstrate how
unintentional method conflicts are untangled in \MIM{}, we only focus on a small subset of the interface model. For example, all the methods declared
in an interface are either default methods or abstract methods. Default methods provide default implementations for methods. Abstract methods do not
have method body, and in fact they can be considered as normal methods that return nothing. Abstract methods can be overridden with future implementations.

\paragraph{Notations}
The metavariables $I, J$ range over interface names; $x$ ranges over variables; $m$ ranges over method names; $e$ ranges over expressions; and $M$ ranges over method declarations. Following Featherweight Java, we assume that the set of variables includes the special variable \kwthis, which cannot be used as the name of an argument to a method. We use the same
conventions as FJ; we write $\overline{I}$ as shorthand for a possibly empty sequence $I_1, ..., I_n$, which may be indexed by $I_i$; and write $\overline{M}$ as shorthand for $M_1 .. M_n$ (with no commas). We also abbreviate operations on pairs of sequences in the obvious way, writing $\overline{I} \; \overline{x}$ for $I_1 \; x_1, ..., I_n \; x_n$, where $n$ is the length of $\overline{I}$ and $\overline{x}$.

\paragraph{Interfaces}
In order to achieve multiple inheritance, an interface can have a set of 
parent interfaces, where such a set can be empty. The interface declaration $\interface{I}{I}{M}$ introduces an interface named $I$ with parent interfaces $\overline{I}$ and a suite of methods $\overline{M}$. The methods of $I$ may either override methods that are already defined in $\overline{I}$ or add new functionality special to $I$, we will illustrate this in more detail later.

\paragraph{Methods}
The method declaration $\method{I}{m}{I_x}{x}{J}{e}$ introduces a
method named $m$ with result type $I$, parameters $\overline{x}$ of
type $\overline{I_x}$ and the overriding target $J$. The body of the
method is the single statement $ \kwreturn \; e;$. We also allow
defining abstract methods with $\absmethod{I}{m}{I_x}{x}{J}$, which is
similar to a normal method but without a method body. Notably, we have introduced the
\kwoverride{} keyword to override an old implementation of the
method. If the interface that it overrides is exactly the enclosing
interface, then such a method is seen as ``originally defined''; otherwise
it is hierarchical overriding.
For simplicity, overloading is not modelled for methods, which
implies that we can uniquely identify a method by its name.

\paragraph{Expressions \& Values}
Expressions can be standard constructs such as variables, method
invocation, or object creation, together with casts. 
Object creation is represented by $\new I$. If any unimplemented abstract methods are defined in interface $I$ (or inherited from parents), then object creation fails. Fields and primitive types are not modelled in \MIM{}. 
Again, to be concise, the casts are merely safe up-casts. Those cast-expressions
are in fact annotated expressions, where the annotation indicates its static type.
The coexistence of static and dynamic types is the key to hierarchical dispatch.
A value
``$(I)\new{J}$''
is the final result of multiple reduction steps evaluating an
expression.

@@

There are two typing rules for method invocation: \textsc{(T-Invk)}
and \textsc{(T-PathInvk)}. \textsc{(T-Invk)} is the typing judgement
for a normal method invocation. First it ensures that the method
receiver $e_0$ type checks to a type $I_0$. Then it checks the type of
the specific method, by using an auxiliary function $\mbody(m, I_0,
I_0)$, which is the function for looking up methods that we will
illustrate in more detail in Section~\ref{subsec:auxdefs}. The form of
\mbody{} is different to FJ, which has two parameters. The first $I_0$
represents the dynamic type and the second, the static type. If the
result of \mbody is not \Undefined, it means that the method exists and the corresponding method parameters and return types are given. After the method type is obtained, we check that all arguments $\overline{e}$ type check to $\overline{I}$ which are subtypes of $\overline{J}$. \textsc{(T-PathInvk)} is the typing judgement for a path invocation. Besides the conditions of \textsc{(T-Invk)}, \textsc{(T-PathInvk)} requires the type of receiver to be the subtype of the specified path type. 
% and \textsc{(T-SuperInvk)} checks if the enclosing type directly extends the specified super type.

@@

Finally, rule \textsc{(T-Anno)} is the typing rule for an expression with static type annotation. The rule says that an annotated expression $\angl{J}e$ type checks to $J$ if the expression $e$ type checks and the static annotation $J$ is a subtype of $I$ which is the type of $e$.
